---
title: "Logical vs. Conditional IF"
author: "Erin"
date: "12/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
```

```{r load libraries}
library(rwebppl)
library(tidyverse)
library(ggrepel)
library(ggthemes)
library(rjson)
char = as.character
num = function(x) {return(as.numeric(char(x)))}
```

The counterfactual in the explanation contains an IF statement:

> IF [premise hadn't happened] THEN [conclusion wouldn't have happened].

This IF could be a conditional statement with a nested inference. In this case, we would sample an actual world and its parameters, then compute counterfactual probabilities for that world in particular.

The IF could also just be a logical IF, that is "Either [premise didn't happen] or [conclusion did happen]". Logical IF is easier to to inference on, since it only involves a single counterfactual sample.

I looked at this before and didn't see much of a difference. Now I'm checking in a very simple, discrete case, where I can enumerate.

```{r load results}
load_results = function(IF) {
  logical_file=paste("../results/ab1d_discrete_", IF, "_A moved B.js", sep="")
  logical_json = fromJSON(
    substring(read_file(logical_file), first=15))
  sapply(logical_json, function(x) {
    return(strsplit(x$label, "\t")[[1]])
  }) %>% t %>% as.data.frame() %>%
    rename(label = V1, endorsement = V2) %>%
    mutate(endorsement = round(num(endorsement), 6),
           IF = IF)
}
# webppl ab1d_discrete_logicalIF.wppl --require ../node_modules/utils --results_dir ../results/ --modules_dir ../node_modules/
df = rbind(load_results("nested_cached"), load_results("nested_no_cache"), load_results("logical_or"), load_results("shallow")) %>%
  spread(IF, endorsement)
```

In the 1D, 2 ball, discrete world prior, logical IF and conditional IF are equivalent.

```{r}
with(df, mean(nested_cached==nested_no_cache))
with(df, mean(nested_cached==shallow))
with(df, mean(nested_cached==logical_or))
```

I'm not sure why the scaling on logical IF using OR is so weird.

```{r, fig.width=5, fig.height=3}
df %>%
  ggplot() +
  aes(x=nested_cached, y=logical_or, label=label) +
  geom_point() +
  geom_text_repel() +
  geom_abline(slope = 1, intercept = 0) +
  theme_few() +
  xlab("Conditional") +
  ylab("!A || B")
ggsave("conditional_vs_logical.png", width=5, height=3)

```


```{r, fig.width=5, fig.height=3}
df %>%
  ggplot() +
  aes(x=nested_cached, y=shallow, label=label) +
  geom_point() +
  geom_text_repel() +
  geom_abline(slope = 1, intercept = 0) +
  theme_few() +
  ylab("Nested Inference") +
  xlab("Shallow Factor Statement")
ggsave("shallow_vs_nested.png", width=5, height=3)
```

shallow counterfactual conditioning takes 0m9.087s
nested coutnerfactual conditioning with caching takes 0m27.053s
nested coutnerfactual conditioning without caching takes 0m26.845s
OR-based IF takes 0m24.242s
