---
title: "Explanations"
author: "Erin Bennett"
header-includes:
   - \usepackage{tikz}
   - \usetikzlibrary{bayesnet}
output: html_document
---

```{r global_options, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = F, warning = F, cache = F, message = F,
                      sanitiz = F, fig.width = 5, fig.height = 3)
```

```{r load_libraries, message=F, warning=F}
source("../utils.R")
```

## 2018 March 11

Speaker shouldn't be inferring semantic variables: doh.

```{r}
speeds = 2
lifted = "lifted"
utterance = "existing"
get_df = function(orig_df) {
  speeds = orig_df$speeds[[1]]
  lifted = orig_df$lifted[[1]]
  utterance = orig_df$utterance[[1]]
  f = paste("../results/s2_exp01_A", utterance,
            "causedBtomove_", lifted, "_",
            speeds, "speeds_s0.53_DNEseparate0.1.json", sep="")
  d = read_file(f) %>%
    paste("[", ., "]") %>%
    fromJSON()
  df = d[1:length(d)%%2==1] %>%
    lapply(as.data.frame) %>%
    do.call(rbind, .) %>%
    mutate(probability = d[1:length(d)%%2==0] %>% unlist()) %>%
    mutate(speeds = speeds, lifted = lifted, utterance = utterance)
  return(df)
}
df = expand.grid(
  speeds = c(2,4,8),
  lifted = c("lifted", "unlifted"),
  utterance = c("moving", "existing")) %>%
  group_by(speeds, lifted, utterance) %>%
  do(get_df(.)) %>%
  ungroup()
```

```{r}
df %>%
  filter(position_label == "AB") %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=w, y=probability, fill=speeds)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(lifted ~ utterance) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 1)
```

```{r}
get_gloss = function(a, b) {
  if (a == -b) {
    return(paste("symmetric collision (", a, ")", sep=""))
  } else if (a == b) {
    if (a > 0) {
      return(paste("A follows B (", a, ")", sep=""))
    } else if (a < 0) {
      return(paste("B follows A (", a, ")", sep=""))
    } else {
      return("no movement")
    }
  } else if (sign(a)==-1 & sign(b)==1) {
    return(paste("opposite directions (", a, ", ", b, ")", sep=""))
  } else if (abs(a) > abs(b)) {
    return(paste("A hits B (", a, ", ", b, ")", sep=""))
  } else {
    return(paste("B hits A (", a, ", ", b, ")", sep=""))
  }
}
get_gloss = function(a, b) {
  if (a == -b) {
    if (a > 0) {
      return("symmetric collision")
    } else if (a < 0) {
      return("opposite directions")
    } else {
      return("no movement")
    }
  } else if (a == b) {
    if (a > 0) {
      return("A follows B")
    } else if (a < 0) {
      return("B follows A")
    } else {
      return("no movement")
    }
  } else if (a==0 & sign(b)==1) {
    return("B moves away from A")
  } else if (b==0 & sign(a)==-1) {
    return("A moves away from B")
  }else if ((sign(a)==-1) & (sign(b)==1)) {
    return("opposite directions")
  } else if (abs(a) > abs(b)) {
    return("A hits B")
  } else {
    return("B hits A")
  }
}
df %>%
  filter(position_label == "AB") %>%
  filter(speeds == 4) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=world, y=probability, 
             # fill=speeds)) +
             colour=gloss)) +
  # geom_bar(stat="identity"#, position="dodge"
  #          ) +
  geom_point() +
  # scale_fill_brewer(type="qual", palette = 6) +
  # scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(lifted ~ utterance, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r, fig.width=15, fig.height=3}
df %>%
  filter(utterance == "existing") %>%
  filter(lifted == "lifted") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity %in% c(-1, 0, 1), b_velocity%in%c(-1, 0, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=speeds, y=probability)) +#, colour=w)) +
  geom_line(aes(group=gloss)) +
  geom_point() +
  # geom_bar(stat="identity", alpha=1/2) +
  facet_grid( ~ gloss, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, lifted, AB")
ggsave("s2_diff_speeds.png", width=15, height=3)
```

```{r, fig.width=5, fig.height=3}
df %>%
  filter(speeds == 8) %>%
  filter(utterance == "existing") %>%
  filter(lifted == "lifted") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity %in% c(-1, 0, 1), b_velocity%in%c(-1, 0, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  # mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=gloss, y=probability)) +#, colour=w)) +
  # geom_line(aes(group=gloss)) +
  # geom_point() +
  geom_bar(stat="identity", alpha=1/2) +
  # facet_grid( ~ gloss, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, lifted, AB")
ggsave("s2.png", width=5, height=3)
```

```{r, fig.width=5, fig.height=3}
df %>%
  filter(speeds == 8) %>%
  filter(utterance == "existing") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity %in% c(-1, 0, 1), b_velocity%in%c(-1, 0, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  # mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=gloss, y=probability, fill=lifted, colour=lifted)) +#, colour=w)) +
  # geom_line(aes(group=gloss)) +
  # geom_point() +
  geom_bar(stat="identity", alpha=1/2, position="dodge") +
  scale_color_solarized() +
  scale_fill_solarized() +
  # facet_grid( ~ gloss, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, lifted, AB")
ggsave("s2_lifted_vs_unlifted_bar.png", width=5, height=3)
```

```{r, fig.width=15, fig.height=5}
df %>%
  filter(utterance == "existing") %>%
  # filter(lifted == "lifted") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity %in% c(-1, 0, 1), b_velocity%in%c(-1, 0, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=speeds, y=probability)) +#, colour=w)) +
  geom_line(aes(group=gloss)) +
  geom_point() +
  # geom_bar(stat="identity", alpha=1/2) +
  facet_grid(lifted ~ gloss, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, AB")
ggsave("s2_diff_speeds_lifted.png", width=15, height=5)
```

```{r, fig.width=15, fig.height=5}
df %>%
  # filter(utterance == "existing") %>%
  filter(lifted == "lifted") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity %in% c(-1, 0, 1), b_velocity%in%c(-1, 0, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, "(", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=speeds, y=probability)) +#, colour=w)) +
  geom_line(aes(group=gloss)) +
  geom_point() +
  # geom_bar(stat="identity", alpha=1/2) +
  facet_grid(utterance ~ gloss, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, lifted, AB")
ggsave("s2_diff_speeds_utterance.png", width=15, height=5)
```

```{r, fig.width=10, fig.height=2.5}
df %>%
  filter(utterance == "existing") %>%
  filter(lifted == "lifted") %>%
  filter(position_label == "AB") %>%
  filter(a_velocity == 1, b_velocity%in%c(-1, -0.5, 0, 0.5, 1)) %>%
  filter(probability != 0) %>%
  mutate(gloss = mapply(get_gloss, a_velocity, b_velocity)) %>%
  mutate(world = paste(gloss, " (", a_velocity, ", ", b_velocity, ")", sep="")) %>%
  mutate(gloss = factor(gloss, levels=c(
    "A hits B",
    "symmetric collision",
    "B hits A",
    "A follows B",
    "B moves away from A",
    "B follows A",
    "A moves away from B",
    "opposite directions"
  ))) %>%
  mutate(world = factor(world, levels=c(
    "symmetric collision (1, -1)",
    "A hits B (1, -0.5)",
    "A hits B (1, 0)",
    "A hits B (1, 0.5)",
    "A follows B (1, 1)"
  ))) %>%
  mutate(w = paste(a_velocity, b_velocity)) %>%
  mutate(speeds = factor(speeds)) %>%
  ggplot(aes(x=speeds, y=probability)) +#, colour=w)) +
  geom_line(aes(group=world)) +
  geom_point() +
  # geom_bar(stat="identity", alpha=1/2) +
  facet_grid(utterance ~ world, scales="free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("s2, existing, lifted, AB")
ggsave("s2_b_velocity_varies.png", width=10, height=2.5)
```

```{r}
speeds = 2
lifted = "lifted"
utterance = "existing"
get_l1 = function(orig_df) {
  speeds = orig_df$speeds[[1]]
  lifted = orig_df$lifted[[1]]
  utterance = orig_df$utterance[[1]]
  f = paste("../results/l1_exp01_A", utterance,
            "causedBtomove_", lifted, "_",
            speeds, "speeds_s0.53_DNEseparate0.1.json", sep="")
  d = read_file(f) %>%
    paste("[", ., "]") %>%
    fromJSON()
  df = d[1:length(d)%%2==1] %>%
    lapply(as.data.frame) %>%
    do.call(rbind, .) %>%
    mutate(probability = d[1:length(d)%%2==0] %>% unlist()) %>%
    mutate(speeds = speeds, lifted = lifted, utterance = utterance)
  return(df)
}
l1 = expand.grid(
  speeds = c(2,4,8),
  lifted = c("lifted", "unlifted"),
  utterance = c("moving", "existing")) %>%
  group_by(speeds, lifted, utterance) %>%
  do(get_l1(.)) %>%
  ungroup()
```

```{r, fig.width=8, fig.height=4}
l1 %>%
  filter(lifted == "lifted") %>%
  # filter(speeds == 4) %>%
  group_by(speeds, lifted, utterance, cf_premise_type, cf_conclusion_type) %>%
  summarise(probability = sum(probability)) %>%
  filter(!is.na(cf_premise_type)) %>%
  ggplot(aes(x=cf_premise_type, y=cf_conclusion_type, fill=probability, label=round(probability, 3))) +
  geom_tile() +
  geom_text() +
  facet_grid(utterance ~ speeds)
```

```{r}
l1 %>%
  filter(!is.na(cf_premise_type)) %>%
  mutate(cf_type = paste(cf_premise_type, cf_conclusion_type)) %>%
  mutate(cf_type = factor(cf_type, levels=c("whether whether", "how whether",
                                            "whether how", "how how"))) %>%
  group_by(utterance, lifted, speeds, w.a_velocity, w.b_velocity, w.position_label) %>%
  mutate(conditional_probability = probability / sum(probability)) %>%
  ungroup() %>%
  # ggplot(aes(x=conditional_probability)) +
  # geom_histogram() +
  # facet_wrap(~cf_type)
  # ggplot(aes(x=cf_type, y=conditional_probability)) +
  ggplot(aes(x=cf_type, y=probability)) +
  stat_summary(fun.data="mean_cl_boot", geom="pointrange")
```



```{r}
l1 %>%
  filter(!is.na(cf_premise_type)) %>%
  mutate(cf_type = paste(cf_premise_type, cf_conclusion_type)) %>%
  mutate(cf_type = factor(cf_type, levels=c(
    "whether whether", "how whether",
    "whether how", "how how"))) %>%
  group_by(utterance, lifted, speeds, w.a_velocity, w.b_velocity, w.position_label) %>%
  mutate(conditional_probability = probability / sum(probability)) %>%
  ungroup() %>%
  # ggplot(aes(x=conditional_probability)) +
  # geom_histogram() +
  # facet_wrap(~cf_type)
  # ggplot(aes(x=cf_type, y=conditional_probability)) +
  ggplot(aes(x=cf_type, y=probability)) +
  facet_wrap(~lifted) + 
  stat_summary(fun.data="mean_cl_boot", geom="pointrange")
```



```{r, fig.width=5, fig.height=3}
l1 %>%
  filter(!is.na(cf_premise_type)) %>%
  mutate(cf_type = paste(cf_premise_type, cf_conclusion_type)) %>%
  mutate(cf_type = factor(cf_type, levels=c("whether whether", "how whether",
                                            "whether how", "how how"))) %>%
  group_by(utterance, lifted, speeds, cf_type, cf_premise_type, cf_conclusion_type) %>%
  summarise(probability = sum(probability)) %>%
  filter(lifted == "lifted", utterance == "existing", speeds==4) %>%
  ggplot(aes(x=cf_premise_type, y=cf_conclusion_type, fill=probability, label=round(probability, 3))) +
  geom_tile() +
  geom_text() +
  # ggplot(aes(x=cf_type, y=probability)) +
  # theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  # facet_wrap(speeds~utterance) +
  # geom_bar(stat="identity") +
  # geom_point() +
  ggtitle("existing, 4, lifted, marginalizing over worlds")
ggsave("how_vs_whether_overall.png", width=5, height=3)
```



specifically, what are the how/whether probabilities for 1 0.5, where B was already moving without A's intervention?

```{r, fig.width=9, fig.height=2.5}
l1 %>% filter(w.a_velocity == 1,
              #w.a_velocity %in% c(-1, 0.5, 1),
              w.b_velocity %in% c(-1, -0.5, 0, 0.5, 1),
              w.position_label == "AB") %>%
  filter(speeds == 4, utterance == "existing", lifted == "lifted") %>%
  ggplot(aes(x=cf_premise_type, y=cf_conclusion_type, fill=probability)) +
  geom_tile() +
  facet_grid(w.a_velocity~w.b_velocity) +
  ggtitle("'How' vs 'whether' as B's velocity varies (A's velocity = 1)")
ggsave("how_vs_whether_a1.0.png", width=9, height=2.5)
```


what is the listener's distribution over worlds for lifted and unlifted?

```{r, fig.width=6, fig.height=3}
l1 %>%
  filter(utterance == "existing", speeds == 4) %>%
  filter(w.position_label == "AB") %>%
  group_by(utterance, speeds, lifted, w.a_velocity, w.b_velocity) %>%
  summarise(probability = sum(probability)) %>%
  ungroup() %>%
  complete(lifted, w.a_velocity, w.b_velocity,
           fill=list(probability=0)) %>%
  ggplot(aes(x=w.a_velocity, y=w.b_velocity, fill=probability)) +
  geom_tile() +
  facet_wrap(~lifted) +
  ggtitle("4, existing")
ggsave("l1_lifted_vs_unlifted.png", width=6, height=3)
```


```{r, fig.width=4, fig.height=3}
l1 %>%
  filter(lifted == "lifted",
         utterance == "existing",
         speeds == 4) %>%
  filter(w.position_label == "AB") %>%
  group_by(utterance, speeds, w.a_velocity, w.b_velocity) %>%
  summarise(probability = sum(probability)) %>%
  ungroup() %>%
  complete(w.a_velocity, w.b_velocity,
           fill=list(probability=0)) %>%
  ggplot(aes(x=w.a_velocity, y=w.b_velocity, fill=probability)) +
  geom_tile() +
  ggtitle("4, existing, lifted")
ggsave("l1.png", width=4, height=3)
```

## 2019 Feb 25

Goal: Reproduce old finding that as the number of discretization bins increases, the endorsement of "A moved B" for the `{A: {velocity: 1, position: 0}, B: {velocity: 0, position: 1}}` world gets higher. All other endorsements go down.

```{r}
parse_label = function(label, var, val) {
  label = strsplit(label, "probability of 'A moved B': ")[[1]][[2]]
  s = strsplit(label, "params: \\{\"probs\":\\[")
  endorsement = num(s[[1]][[1]])
  label = s[[1]][[2]]
  probs = strsplit(strsplit(label, "\\]")[[1]][[1]], ",")[[1]] %>% num()
  new_var = c("endorsement", "ww", "wh", "hw", "hh")
  new_val = c(endorsement, probs)
  data.frame(var=c(var, new_var), val=c(val, new_val)) %>%
    return()
}
  results_dir = function(f) {
    return(paste("~/Projects/causal_simulations/results/2019jan/", f, sep=""))
  }
read_explanations_rs = function(n) {
  f = results_dir(paste("s1_enumeration_AmovedB_", n, "velocities_s0.53.js", sep=""))
  d = read_file(f) %>%
    substring(14) %>%
    fromJSON()
  df = d$world$A %>%
    mutate(ball="A",
           label=d$label,
           id=1:length(ball)) %>%
    rbind(d$world$B %>%
            mutate(ball="B",
                   label=d$label,
                   id=1:length(ball))) %>%
    gather("var", "val", c(velocity, position)) %>%
    unite("var", ball, var)
  df %>% group_by(id) %>%
    do(parse_label(.$label, .$var, .$val)) %>%
    ungroup() %>%
    spread(var, val) %>%
    mutate(n = n) %>%
    return()
}
get_world_name = function(a, b, c, d) {
  if (a==-1 & b==-1) {
    return("B follows A")
  } else if (a==1 & b==1) {
    return("A follows B")
  } else if (a==1 & b==0) {
    return("A hits B")
  } else if (a==0 & b==-1) {
    return("B hits A")
  } else if (a==1 & b==-1) {
    return("symmetric collision")
  } else {
    return("NA")
  }
}
df_old = do.call(rbind, lapply(c(3,5,7,11), read_explanations_rs)) %>%
  mutate(n = factor(n)) %>%
  mutate(name = mapply(get_world_name, A_velocity, B_velocity, A_position, B_position))
```

```{r, fig.width=8, fig.height=2}
df_old %>%
  ggplot(aes(x=n, y=endorsement)) +
  geom_bar(stat="identity", alpha=1/2) +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(~name) +
  ylim(0, 1)
# ggsave(results_dir("endorsement_by_velocity.png"), width=10, height=3)
```


```{r}
# parse_label = function(label, var, val) {
#   label = strsplit(label, "probability of 'A moved B': ")[[1]][[2]]
#   s = strsplit(label, "params: \\{\"probs\":\\[")
#   endorsement = num(s[[1]][[1]])
#   label = s[[1]][[2]]
#   probs = strsplit(strsplit(label, "\\]")[[1]][[1]], ",")[[1]] %>% num()
#   new_var = c("endorsement", "ww", "wh", "hw", "hh")
#   new_val = c(endorsement, probs)
#   data.frame(var=c(var, new_var), val=c(val, new_val)) %>%
#     return()
# }
#   results_dir = function(f) {
#     return(paste("~/Projects/causal_simulations/results/", f, sep=""))
#   }
# read_explanations_rs = function(n) {
#   f = results_dir(paste("s1_enumeration_AmovedB_", n, "velocities_s0.53.js", sep=""))
#   d = read_file(f) %>%
#     substring(14) %>%
#     fromJSON()
#   df = d$world$A %>%
#     mutate(ball="A",
#            label=d$label,
#            id=1:length(ball)) %>%
#     rbind(d$world$B %>%
#             mutate(ball="B",
#                    label=d$label,
#                    id=1:length(ball))) %>%
#     gather("var", "val", c(velocity, position)) %>%
#     unite("var", ball, var)
#   df %>% group_by(id) %>%
#     do(parse_label(.$label, .$var, .$val)) %>%
#     ungroup() %>%
#     spread(var, val) %>%
#     mutate(n = n) %>%
#     return()
# }
# get_world_name = function(a, b, c, d) {
#   if (a==-1 & b==-1) {
#     return("B follows A")
#   } else if (a==1 & b==1) {
#     return("A follows B")
#   } else if (a==1 & b==0) {
#     return("A hits B")
#   } else if (a==0 & b==-1) {
#     return("B hits A")
#   } else if (a==1 & b==-1) {
#     return("symmetric collision")
#   } else {
#     return("NA")
#   }
# }
# df_rep = do.call(rbind, lapply(c(3,5#,7,11
#                              ), read_explanations_rs)) %>%
#   mutate(n = factor(n)) %>%
#   mutate(name = mapply(get_world_name, A_velocity, B_velocity, A_position, B_position))
# approx_eq((df_rep %>% filter(n %in% c(3,5)) %>% .$endorsement),
#     (df_old %>% filter(n %in% c(3,5)) %>% .$endorsement)) %>% mean()==1
```


```{r}
parse_label = function(label, var, val) {
  data_parts = strsplit(label, "<br/>probability of '.*': ")[[1]]
  gloss = strsplit(data_parts[[1]], "gloss: ")[[1]][[2]]
  endorsement = num(data_parts[[2]])
  # label = strsplit(label, "probability of 'A moved B': ")[[1]][[2]]
  # s = strsplit(label, "params: \\{\"probs\":\\[")
  # endorsement = num(s[[1]][[1]])
  # label = s[[1]][[2]]
  # probs = strsplit(strsplit(label, "\\]")[[1]][[1]], ",")[[1]] %>% num()
  # new_var = c("endorsement", "ww", "wh", "hw", "hh")
  # new_val = c(endorsement, probs)
  new_var = c("gloss", "endorsement")
  new_val = c(gloss, endorsement)
  data.frame(var=c(var, new_var), val=c(val, new_val)) %>%
    return()
}
  results_dir = function(f) {
    return(paste("~/Projects/causal_simulations/results/", f, sep=""))
  }
read_explanations_rs = function(fname) {
  f = results_dir(fname)
  d = read_file(f) %>%
    substring(14) %>%
    fromJSON()
  df = d$world$A %>%
    mutate(ball="A",
           label=d$label,
           id=1:length(ball)) %>%
    rbind(d$world$B %>%
            mutate(ball="B",
                   label=d$label,
                   id=1:length(ball))) %>%
    gather("var", "val", c(velocity, position)) %>%
    unite("var", ball, var)
  df %>% group_by(id) %>%
    mutate(var = char(var), val = char(val)) %>%
    do(parse_label(.$label, .$var, .$val)) %>%
    ungroup() %>%
    spread(var, val) %>%
    mutate(A_position = num(A_position),
           B_position = num(B_position),
           A_velocity = num(A_velocity),
           B_velocity = num(B_velocity),
           endorsement = num(endorsement)) %>%
    mutate(fname = fname) %>%
    separate(fname, c("model", "prior", "utterance", "lifted", "nspeeds", "ignore", "stickiness_percent", "dne", "disjoint")) %>%
    return()
}

df = do.call(rbind, lapply(c(
  "s1_exp01_AmovingcausedBtomove_lifted_1speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_lifted_2speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_lifted_3speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_lifted_4speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_lifted_5speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_lifted_8speeds_s0.53_noDNE.js",
  
  "s1_exp01_AmovingcausedBtomove_unlifted_1speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_3speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_noDNE.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_5speeds_s0.53_noDNE.js",
  
  "s1_exp01_AmovingcausedBtomove_lifted_1speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovingcausedBtomove_lifted_2speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovingcausedBtomove_lifted_3speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovingcausedBtomove_lifted_4speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovingcausedBtomove_lifted_5speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovingcausedBtomove_lifted_8speeds_s0.53_DNEisvelocity.js",
  
  "s1_exp01_AmovedB_lifted_1speeds_s0.53_DNEisvelocity.js",
  # "s1_exp01_AmovedB_lifted_2speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_lifted_3speeds_s0.53_DNEisvelocity.js",
  # "s1_exp01_AmovedB_lifted_4speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_lifted_5speeds_s0.53_DNEisvelocity.js",
  # "s1_exp01_AmovedB_lifted_8speeds_s0.53_DNEisvelocity.js",
  
  "s1_exp01_AmovedB_unlifted_1speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_unlifted_2speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_unlifted_3speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_unlifted_4speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_unlifted_5speeds_s0.53_DNEisvelocity.js",
  
  "s1_exp01_AmovedB_lifted_2speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovedB_lifted_4speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovedB_lifted_8speeds_s0.53_DNEseparate.js",
  
  "s1_exp01_AmovedB_lifted_2speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_lifted_4speeds_s0.53_DNEisvelocity.js",
  "s1_exp01_AmovedB_lifted_8speeds_s0.53_DNEisvelocity.js",
  
  "s1_exp01_AmovedB_unlifted_2speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovedB_unlifted_4speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovedB_unlifted_8speeds_s0.53_DNEseparate.js",
  
  "s1_exp01_AmovingcausedBtomove_lifted_2speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovingcausedBtomove_lifted_4speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovingcausedBtomove_lifted_8speeds_s0.53_DNEseparate.js",
  
  "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_DNEseparate.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_8speeds_s0.53_DNEseparate.js",
  
  "s1_exp01_AmovingcausedBtomove_lifted_2speeds_s0.53_DNEseparate0.1.js",
  "s1_exp01_AmovingcausedBtomove_lifted_4speeds_s0.53_DNEseparate0.1.js",
  
  "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_DNEseparate0.1.js",
  "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_DNEseparate0.1.js"
  ), read_explanations_rs))
```

```{r}
# approx_eq(
#   (df %>% filter(endorsement > 0.5) %>% filter(nspeeds == "1speeds") %>%
#      .$endorsement),
#   (df_old %>% filter(n %in% c(3)) %>% .$endorsement)) %>% mean()==1
```

```{r, fig.width=8, fig.height=4}
df %>%
  mutate(params = paste(dne, utterance)) %>%
  filter(A_velocity %in% c(-1, 0, 1)) %>%
  filter(B_velocity %in% c(-1, 0, 1)) %>%
  # filter(nspeeds %in% paste(c(2,4,8), "speeds", sep="")) %>%
  mutate(utterance=ifelse(utterance=="AmovedB", "AexistingcausedBtomove",
                          char(utterance))) %>%
  filter(endorsement != 0) %>%
  ggplot(aes(x=gloss, y=endorsement, fill=nspeeds, colour=nspeeds)) +
  geom_hline(yintercept = 0.5, linetype="dashed", alpha=1/2) +
  geom_bar(stat="identity", position="dodge"#, alpha=1/2
           ) +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(lifted~params) +
  # facet_grid(lifted~params) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 1)
ggsave(results_dir("endorsement_by_velocity.png"), width=20, height=8)
```

Successful!

### Explanation

As of commit `932ce49`, the `--lifted` parameter wasn't actually lifting the inference of the semantic variables (`cf_premise_type` and `cf_conclusion_type`) up to the S1 level. And it turns out this matters. Commit `b6d26e2` fixes this bug.

### Notes for next time I pick this up

- position discretization would be nice to explore.
- difference between Amoving premise and Aexisting premise is that Aexisting allows endorsement for worlds where B hits A to be more than 0.
- unpack how/whether, especially for slow-moving B being hit by A.


```{r, fig.width=8, fig.height=4}
df %>%
  filter(utterance == "AmovingcausedBtomove") %>%
  filter(dne == "DNEseparate0") %>%
  filter(lifted == "lifted") %>%
  filter(endorsement != 0) %>%
  filter(A_velocity %in% c(-1, 0, 1)) %>%
  filter(B_velocity %in% c(-1, 0, 1)) %>%
  filter(nspeeds %in% paste(c(2,4,8), "speeds", sep="")) %>%
  ggplot(aes(x=gloss, y=endorsement, fill=nspeeds, colour=nspeeds)) +
  geom_hline(yintercept = 0.5, linetype="dashed", alpha=1/2) +
  geom_bar(stat="identity", position="dodge", alpha=1/2) +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(~lifted) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 1)
ggsave(results_dir("velocity.png"), width=12, height=5)
```


```{r, fig.width=8, fig.height=4}
df %>%
  filter(utterance == "AmovingcausedBtomove") %>%
  filter(dne == "DNEseparate0") %>%
  filter(endorsement != 0) %>%
  filter(A_velocity %in% c(-1, 0, 1)) %>%
  filter(B_velocity %in% c(-1, 0, 1)) %>%
  filter(nspeeds %in% paste(c(2,4,8), "speeds", sep="")) %>%
  ggplot(aes(x=gloss, y=endorsement, fill=nspeeds, colour=nspeeds)) +
  geom_hline(yintercept = 0.5, linetype="dashed", alpha=1/2) +
  geom_bar(stat="identity", position="dodge", alpha=1/2) +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(~lifted) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 1)
ggsave(results_dir("liftedness.png"), width=12, height=5)
```

```{r, fig.width=8, fig.height=4}
df %>%
  filter(utterance == "AmovingcausedBtomove") %>%
  filter(lifted == "lifted") %>%
  filter(endorsement != 0) %>%
  filter(A_velocity %in% c(-1, 0, 1)) %>%
  filter(B_velocity %in% c(-1, 0, 1)) %>%
  # filter(nspeeds == "2speeds") %>%
  filter(nspeeds %in% c("2speeds", "4speeds")) %>%
  mutate(dne = factor(
    dne,
    levels=c("noDNE", "DNEseparate0", "DNEseparate", "DNEisvelocity"),
    labels=c("Always exist", "Constant probability (0.1) of not existing", "Constant probability (0.5) of not existing", "Nonexistence is a 'velocity' state"))) %>%
  mutate(utterance=ifelse(utterance=="AmovedB", "AexistingcausedBtomove",
                          char(utterance))) %>%
  # filter(endorsement != 0) %>%
  # ggplot(aes(x=gloss, y=endorsement, fill=paste(A_velocity, B_velocity))) +
  ggplot(aes(x=gloss, y=endorsement, fill=nspeeds, colour=nspeeds)) +
  geom_hline(yintercept = 0.5, linetype="dashed", alpha=1/2) +
  geom_bar(stat="identity", position="dodge", alpha=1/2) +
  # geom_line(aes(group=dne), alpha=1/2, colour="black") +
  # geom_point(#position=position_dodge(0.5), 
    #alpha=1/2, 
    # size=2) +
  scale_fill_brewer(type="qual", palette = 6) +
  scale_colour_brewer(type="qual", palette = 6) +
  facet_grid(~dne) +
  # facet_grid(lifted~params) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("\"A moving caused B to move\"") +
  ylim(0, 1)
ggsave(results_dir("dne.png"), width=16, height=6)
```

```{r}
fname = "archive/l0_exp01_AmovedB_false_false_unlifted_4speeds_s0.53_DNEseparate.js"
fpath = paste("../results/", fname, sep="")
d = read_file(fpath) %>%
  substring(14) %>%
  fromJSON()
df = data.frame(
  A_velocity = d$world$A$velocity,
  A_position = d$world$A$position,
  B_velocity = d$world$B$velocity,
  B_position = d$world$B$position,
  probability = num(d$label))
df %>%
  filter(A_position==0 & B_position==1) %>%
  select(A_velocity, B_velocity, probability) %>%
  rbind(data.frame(A_velocity=0, B_velocity=0, probability = 0)) %>%
  complete(A_velocity, B_velocity, fill=list(probability=0)) %>%
  ggplot(aes(x=A_velocity, y=B_velocity, fill=probability)) +
  geom_tile()
ggsave("../results/l0.png", width=8, height=5)
```


```{r}




  df = d$world$A %>%
    mutate(ball="A",
           label=d$label,
           id=1:length(ball)) %>%
    rbind(d$world$B %>%
            mutate(ball="B",
                   label=d$label,
                   id=1:length(ball))) %>%
    gather("var", "val", c(velocity, position)) %>%
    unite("var", ball, var)
  df %>% group_by(id) %>%
    mutate(var = char(var), val = char(val)) %>%
    do(parse_label(.$label, .$var, .$val)) %>%
    ungroup() %>%
    spread(var, val) %>%
    mutate(A_position = num(A_position),
           B_position = num(B_position),
           A_velocity = num(A_velocity),
           B_velocity = num(B_velocity),
           endorsement = num(endorsement)) %>%
    mutate(fname = fname) %>%
    separate(fname, c("model", "prior", "utterance", "lifted", "nspeeds", "ignore", "stickiness_percent", "dne", "disjoint")) %>%
    return()
```









```{r}
# parse_label = function(label, var, val) {
#   # data_parts = strsplit(label, "<br/>probability of '.*': ")[[1]]
#   # gloss = strsplit(data_parts[[1]], "gloss: ")[[1]][[2]]
#   # endorsement = num(data_parts[[2]])
#   label = strsplit(label, "probability of 'A moved B': ")[[1]][[2]]
#   s = strsplit(label, "params: \\{\"probs\":\\[")
#   endorsement = num(s[[1]][[1]])
#   label = s[[1]][[2]]
#   probs = strsplit(strsplit(label, "\\]")[[1]][[1]], ",")[[1]] %>% num()
#   new_var = c("endorsement", "ww", "wh", "hw", "hh")
#   new_val = c(endorsement, probs)
#   # new_var = c("gloss", "endorsement")
#   # new_val = c(gloss, endorsement)
#   data.frame(var=c(var, new_var), val=c(val, new_val)) %>%
#     return()
# }
#   results_dir = function(f) {
#     return(paste("~/Projects/causal_simulations/results/", f, sep=""))
#   }
# read_explanations_rs = function(fname) {
#   f = results_dir(fname)
#   d = read_file(f) %>%
#     substring(14) %>%
#     fromJSON()
#   df = d$world$A %>%
#     mutate(ball="A",
#            label=d$label,
#            id=1:length(ball)) %>%
#     rbind(d$world$B %>%
#             mutate(ball="B",
#                    label=d$label,
#                    id=1:length(ball))) %>%
#     gather("var", "val", c(velocity, position)) %>%
#     unite("var", ball, var)
#   df %>% group_by(id) %>%
#     mutate(var = char(var), val = char(val)) %>%
#     do(parse_label(.$label, .$var, .$val)) %>%
#     ungroup() %>%
#     spread(var, val) %>%
#     mutate(A_position = num(A_position),
#            B_position = num(B_position),
#            A_velocity = num(A_velocity),
#            B_velocity = num(B_velocity),
#            endorsement = num(endorsement)) %>%
#     mutate(fname = fname) %>%
#     separate(fname, c("model", "prior", "utterance", "lifted", "nspeeds", "ignore", "stickiness_percent", "dne", "disjoint")) %>%
#     return()
# }
```


```{r}
# df = do.call(rbind, lapply(c(
#   "s1_enumeration_AmovedB_3velocities_s0.53.js"
#   # "s1_exp01_AmovingcausedBtomove_unlifted_1speeds_s0.53_noDNE.js"
#   # "s1_exp01_AmovingcausedBtomove_unlifted_1speeds_s0.53_noDNE.js",
#   # "s1_exp01_AmovingcausedBtomove_lifted_1speeds_s0.53_noDNE.js",
#   # "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_noDNE.js",
#   # "s1_exp01_AmovingcausedBtomove_lifted_2speeds_s0.53_noDNE.js"
#   # "s1_exp01_AmovingcausedBtomove_unlifted_1speeds_s0.53_noDNE.js",
#   # "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_noDNE.js"
#   # # "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_noDNE.js",
#   # # "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_noDNE.js",
#   # "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_DNEisvelocity.js",
#   # "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_DNEisvelocity.js",
#   # "s1_exp01_AmovedB_unlifted_2speeds_s0.53_DNEisvelocity.js",
#   # "s1_exp01_AmovedB_unlifted_4speeds_s0.53_DNEisvelocity.js"
#   # # "s1_exp01_AmovingcausedBtomove_unlifted_2speeds_s0.53_DNEseparate.js",
#   # # "s1_exp01_AmovingcausedBtomove_unlifted_4speeds_s0.53_DNEseparate.js"
#   ), read_explanations_rs))
```

```{r, fig.height=2, fig.width=5}
# df %>%
#   ggplot(aes(x=paste(A_velocity, B_velocity),#gloss,#paste(gloss, A_velocity, B_velocity),
#              fill=lifted,#fname,#paste(nspeeds, dne, lifted),
#              y=endorsement)) +
#   geom_bar(stat="identity", position="dodge") +
#   facet_wrap(stickiness_percent~utterance) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#   geom_hline(yintercept = 0.5, linetype="dashed")
```


```{r, fig.width=10, fig.height=5}
# df %>%
#   ggplot(aes(x=A_velocity, y=B_velocity,
#              colour=endorsement, fill=endorsement)) +
#   geom_tile() +
#   facet_wrap(~paste(utterance, nspeeds, dne))
```


